A template for the *arxiv* style

### 2022-02-07 4-5-interactive.Rmd
small changes in read-through

With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable or to find patterns such as trend or seasonality in the time series. Combining this two types of plot with interactivity let users to link between points on the map and their corresponding time series and explore the spatial and temporal dimension of the data simultaneously. Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk}.

Starting with the original data, some pre-processing may be required to summarise the data before the visualisation. This example explores the variation of monthly temperature range across `r nrow(weatherdata::climate_full)` weather stations in Australia and with `weatherdata::climate_full`, daily temperature range is first calculated as the difference between \code{tmax} and \code{tmin}. The three daily variables are then averaged into month over 2016 - 2020 in the long form. Variance of the temperature difference is then calculated for each station in the nested form. Then, a \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column. The spatial map and time series plot are then made with each \code{SharedData} objects separately. In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance and a ribbon band is constructed using the long form cubble to show the maximum and minimum temperature of each station across month. With a different dataset, users are free to calculate any per station measure in the nested form or to make any time-wise summary of the data in the long form to customise the spatial or temporal view. And the cross-linking is always safeguarded by the shared \code{id} column embedded in the cubble structure. Below is the pseudo code that outlines the process to construct an interactive graphic described above:

@@

In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic. Overall, most regions in Australia have low variance of temperature range across different months while the north-west coastline, bottom of South Australia, and Victoria stands out with larger monthly changes. In the second row, Mount Elizabeth is selected on the map given its dark colour and this links to the ribbon on the right where there is a larger temperature range is presented in Australian winter (June to August). The third row selects the Grampian station in Victoria and the linked ribbon shows an opposite wider range in Australian summer period (December to February). The last row selects point with the lowest minimum temperature in August in the ribbon plot and surprisingly, this links to the Thredbo Airport in the Victoria and New South Wales border, rather than somewhere in the Tasmania island!

@@

With leaflet popup, the temperature range can be displayed as a small subplot upon clicking on the map. This would require first creating the popup plots from the long form cubble as a vector and then add these plots to a leaflet map, created from the nested cubble, with \code{leafpop::addPopupGraphs()}:

@@

Figure \ref{fig:interactive-popup} shows the same content as Figure \ref{fig:interactive-linking} but made with leaflet and popups.

```{r interactive-popup, echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as FIgure 11 with thetemperature variation shown as a popup in the leaflet map.", fig.show='hold'}
knitr::include_graphics(here::here("figures/popup-mid.png")) 
```

>>>>

With spatio-temporal data, users may wish to make plots to learn the spatial distribution of a variable, or to find patterns, such as trend or seasonality, in the time series. Combining this two types of plot with interactivity let users to link between points on the map and the corresponding time series to explore the spatial and temporal dimension of the data simultaneously. Below is an example that describes the process of building an interactive graphic with \pkg{cubble} and \pkg{crosstalk} The example explores the variation of monthly temperature range with \code{weatherdata::climate_full} data.

The temperature range is calculated as the difference between \code{tmax} and \code{tmin} and its monthly average over 2016 - 2020 is taken before calculating the variance. A \code{SharedData} object is constructed for each form of the cubble and the same \code{group} argument ensures the cross-linking of the two forms via the common \code{id} column. The spatial map and time series plot are then made with each \code{SharedData} objects separately. In this example, stations on the Australia map, made from the nested form, are coloured by the calculated variance and a ribbon band is constructed using the long form cubble to show the maximum and minimum temperature of each station across month. With a different dataset, users are free to calculate any per station measure in the nested form or to make any time-wise summary of the data in the long form to customise the spatial or temporal view. The cross-linking between the two plots is always safeguarded by the shared \code{id} column embedded in the cubble structure. Below is the pseudo code that outlines the process to construct an interactive graphic described above:

@@

In Figure \ref{fig:interactive-linking}, the first row shows the initial view of the interactive graphic. On the map, most regions in Australia have low variance of temperature range while the north-west coastline, bottom of South Australia, and Victoria stands out with larger monthly changes. In the second row, Mount Elizabeth is selected on the map given its high variance colour on the initial map and this links to the ribbon on the right. The third row the lowest temperature in August and this corresponds to Thredbo AWS in the Victoria and New South Wales border. Another station in the Tasmania island is selected on the map to cross compare with Thredbo AWS.

@@

This plot can also be made using \pkg{cubble} and \pkg{leaflet} where the temperature range can be displayed as a small subplot upon clicking on the map. This would require first creating the popup plots from the long form cubble as a vector and then add these plots to a leaflet map created from the nested cubble, with \code{leafpop::addPopupGraphs()}:

@@

Figure \ref{fig:interactive-popup} shows Figure \ref{fig:interactive-linking} made made with leaflet and popups.

```{r interactive-popup, echo = FALSE, out.width="45%", out.height="25%", fig.retina = 2, dpi = 300, fig.cap = "Same as Figure 11 with thetemperature variation shown as a popup in the leaflet map.", fig.show='hold'}
knitr::include_graphics(here::here("figures/popup-mid.png")) 
```

### 2022-02-07 5-conclusion.Rmd
small changes in read-through

This paper describes the \proglang{R} package \pkg{cubble} for manipulating and visualising spatio-temporal data. A new data structure, that builds from the \code{rowwise_df} and \code{grouped_df} class in the tidyverse ecosystem, is proposed to connect the spatial and temporal variables in the spatio-temporal data. This design frees the data analysts from spending time on organising variables of different observational units. The data structure is also flexible to the techniques and packages analysts use to analyse the data, for example, in the matching example in section 4.3, users are free to use algorithms from another package to cluster stations. Further development and maintenance of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}. Another area for further development is to extend \pkg{cubble} to other spatial objects other than points.

>>>>

This paper describes an \proglang{R} package \pkg{cubble} for manipulating and visualising spatio-temporal data. A new data structure, \code{cubble} that builds from the \code{rowwise_df} and \code{grouped_df} class in the tidyverse ecosystem, is proposed to connect the time invariant and varying variables in the spatio-temporal data. This design frees the data analysts from spending time on organising variables of different observational units. The data structure is also flexible to the techniques and packages analysts use to analyse the data, for example, in the matching example in section 4.3, users are free to use algorithms from another package to cluster stations.

Further development and maintenance of the package involves responding to changes in the tidyverse packages that \pkg{cubble} imports, in particular, \pkg{tibble}, \pkg{tidyr}, and \pkg{dplyr}. Another area for further development is to extend \pkg{cubble} to other spatial objects other than points.

### 2022-02-07 main.Rmd
small changes in read-through

---
documentclass: jss
author:
  - name: H. Sherry Zhang
    affiliation: Monash University
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email: \email{huize.zhang@monash.edu}
  - name: Dianne Cook
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    # location: Melbourne, Australia
    email:  dicook@monash.edu
  - name: Ursula Laa
    affiliation: 'University of Natural Resources and Life Sciences \AND'
    address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
    email:  ursula.laa@boku.ac.at
  - name: Nicolas Langrené
    affiliation: 'BNU-HKBU United International College \AND'
    address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
    email: nicolas@langrene.com
  - name: Patricia Menéndez
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email:  patricia.menendez@monash.edu
title:
  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
  # If you use tex in the formatted title, alsoge supply version without
  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
  # For running headers, if needed
  # short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  Sptio-temporal data naturally has two levels of observations: site and time level. [problem]. In this paper, a new data structure, `cubble`, is proposed for manipulating and visualising spatio-temporal data.
keywords:
  # at least one keyword must be supplied
  formatted: [spatio temporal,  "\\proglang{R}", climate weather station, data analysis]
  plain:     [spatio temporal, R, climate weather station, data analysis]
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
  \usepackage{xcolor}
output: rticles::jss_article
bibliography: references.bib
---

>>>>

---
documentclass: jss
author:
  - name: H. Sherry Zhang
    affiliation: Monash University
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email: \email{huize.zhang@monash.edu}
  - name: Dianne Cook
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    # location: Melbourne, Australia
    email:  dicook@monash.edu
  - name: Ursula Laa
    affiliation: 'University of Natural Resources and Life Sciences \AND'
    address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
    email:  ursula.laa@boku.ac.at
  - name: Nicolas Langrené
    affiliation: 'BNU-HKBU United International College \AND'
    address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
    email: nicolas@langrene.com
  - name: Patricia Menéndez
    affiliation: 'Monash University \AND'
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email:  patricia.menendez@monash.edu
title:
  formatted: "\\pkg{cubble}: An R Package for Structuring Spatio-temporal Data"
  # If you use tex in the formatted title, alsoge supply version without
  plain:     "cubble: An R Package for Structuring Spatio-temporal Data"
  # For running headers, if needed
  # short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  Spatio-temporal variables can be divided into time invariant ones that can be identified by the spatial identifier, or the time varying ones that cross-identified by the spatial and temporal identifiers. In this paper, a new data structure, \code{cubble}, is proposed for manipulating and visualising spatio-temporal data. The new data structure uses two forms to organise time invariant/ varying variables so that the spatial and temporal dimension of the data can be manipulated separately while keep synchronised. Advanced considerations are given to illustrate data with hierarchical structure, data matching, and how cubble fits with the interactive graphics pipeline. Examples are given to analysing Australia meteorology data, river level data, and climate reanalysis (ERA5). 
keywords:
  # at least one keyword must be supplied
  formatted: [spatio temporal,  "\\proglang{R}", climate weather station, data analysis]
  plain:     [spatio temporal, R, climate weather station, data analysis]
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
  \usepackage{xcolor}
output: rticles::jss_article
bibliography: references.bib
---
